### 2023-12-29 tesis_Ashish/Chapter7/chapter7.tex
Updated thesis 29 Dec 2023

The Phase II upgrade of the CMS Tracker is a crucial project undertaken to maintain the effectiveness of the CMS experiment during the High Luminosity LHC (HL-LHC) phase, which is expected to start in 2029. The HL-LHC will significantly increase the instantaneous luminosity, leading to a higher number of proton-proton collisions per bunch crossing. Instantaneous luminosity will increase to an unprecedented value of $7.5 \times 10^{34} cm^{-2} \: s^{-1}$ which corresponds to 200 proton-proton collisions per bunch crossing. This higher collision rate poses several challenges, including increased radiation levels, higher data rates, and more complex event environment. Run 2 pixel detector will not be able to handle these extreme radiation environment, resolve nearby particle tracks and operate properly to give a reliable estimate of the instantaneous luminosity for such a high pileup. That is why it will be replaced by a new pixel detector. The Phase II upgrade of the CMS tracker comprises of two main components: the Inner Tracker (IT) and the Outer Tracker (OT) \cite{collaboration:2759074}. The Inner Tracker consists of three main sub-components: the Tracker Endcap Pixel detector (TEPX), the Tracker Forward Pixel detector (TFPX), and the Tracker Barrel Pixel (TBPX) detector as shown in Fig. \ref{fig:Innertracker}.  %The Outer Tracker is responsible for detecting charged particles in the outer region of the CMS detector. It consists of two key subsystems: the silicon strip tracker and the silicon macro pixel tracker \cite{RoyChowdhury:2729279}.

@@

TEPX Disk 4 Ring 1 (D4R1) is located in the innermost ring of the last double-disk of the TEPX, is uniquely designed for measuring luminosity and beam-induced background (BIB). D4R1 is optimally placed due to its low radius, large active area, and high z position, making it highly sensitive and efficient for these measurements. The D4R1 operational model is distinct from the rest of TEPX. It employs the entire available trigger rate and readout bandwidth, which varies based on operational mode (binary readout or Time-over-Threshold information). This capability ranges from a trigger rate of 825 kHz (similar to the rest of TEPX) up to several MHz, limited by readout bandwidth and the available e-link capacity. In nominal physics conditions, the baseline trigger rate for D4R1 is 825 kHz, while in van der Meer (vdM) conditions, it is 2 MHz, with the potential to reach between 5 and 10 MHz in certain scenarios.

@@

Fig. \ref{fig:tepx_cl} shows the location of TEPX clusters in Disk 4 in the XY plane (Z is the direction of the beam) from a simulated sample with a pileup of 200. %Clusters refer to groups of hits in the pixel detector.
It consists of five concentric rings that are color-coded to indicate the density of clusters within the detector. It is transitioning from blue (low density) to red (high density), with the hottest regions indicating the overlap modules within the rings. The granularity of the data is high, with the number of bins set to 1000 for both axes and a bin size of 0.1 cm, allowing for a detailed visualization of the clusters across the disk.
%The TEPX disk in this map consists of 5 rings which are concentric rings of pixel modules at different radial distances from the beamline.

@@

Simlarly, we also studied two-fold coincidences in r which arise from module overlaps between successive rings by only considering one to one module overlap between first two and last two layers in one TEPX double disk.  %Table \ref{table:two_fold_coincidences} categorizes various types of two-fold coincidences in r, observed in the TEPX detector. These coincidences are labeled from A to L and arise due to the module overlap in different layers within a single TEPX double disk. The 'dz' column quantifies the spatial separation along beam axis between the coinciding layers,  the z coordinates for Layers L1 to L4 within the TEPX detector is also mentioned. These coordinates provide a precise location for each layer, contributing to the understanding of where these two fold coincidences are likely to occur. The two fold coincidence in r map showing module overlap regions for different types is shown in Fig. \ref{fig:cluster_ring_770}.
 A one-to-one overlapping of modules is considered between L1L2 and L3L4 module overlap as shown in Fig. \ref{fig:cluster_ring_72034}. %as shown in Fig. \ref{tab:my_label2}. 
 %The distributions for dr, dphi distribution depicting fake, true, total two fold coincidence in r for pileup 0.5 and 200 are shown in Fig. \ref{fig:cluster_ring_81} and \ref{fig:cluster_ring_82} respectively.
 Fit is done to pileup 0.5 distribution of true coincidences for both L1L2 and L3L4 coincidences in r. %as shown in Fig. \ref{fig:cluster_ring_83} and \ref{fig:cluster_ring_84}. 
  Fit results for all disks and rings are shown from  Table \ref{tab:my_label_55} to \ref{tab:my_label_54}. Similarly, two-sigma standard deviation selections are applied on dr, dphi variables. After applying all selections, the XY map of two-fold coincidences in r for L1L2 and L3L4 combined as shown in \ref{fig:cluster_ring_79}.

@@

\begin{table}[htbp]
  \centering
  \caption[Expected stat. precision of TEPX under low pileup] {Expected statistical precision (in $\%$) in head-on collisions during typical vdM conditions with pileup of 0.5 for TEPX clusters and two-fold coincidences over different integration time period.}
\begin{tabular}{lcccc}
vdM (PU 0.5) & Trigger Rate (kHz) & 1 bx, 1s & 1 bx, 30s\\
\hline
TEPXD4R1 Clusters&1000&1.82&0.332\\
TEPXD4R1 2x Coincidences in phi &1000&5.98&1.09\\
TEPXD4R1 2x Coincidences in r &1000&11.1&2.02\\
TEPXD4R1 2x Coincidences &1000&5.27&0.962\\
TEPX Clusters&500&0.709&0.129\\
TEPX 2x Coincidences in phi &500&2.65&0.485\\
TEPX 2x Coincidences in r &500&4.59&0.838\\
TEPX 2x Coincidences &500&2.3&0.42\\
\end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\caption[Expected stat. precision of TEPX under high pileup]{Expected statistical precision (in $\%$) in head-on collisions during physics conditions with pileup of 200 for TEPX clusters and two-fold coincidences over different integration time period.}
\begin{small} % this command will decrease the font size
\begin{tabular}{lcccc}
Algorithm & Trigger Rate (kHz) & 1 bx, 1s & 2500 bx, 1s\\
\hline
TEPXD4R1 Clusters & 825 & 0.1 & 0.002  \\
TEPXD4R1 2x Coincidences in phi & 825 & 0.329 & 0.00659 \\
TEPXD4R1 2x Coincidences in r & 825 & 0.61 & 0.0122  \\
TEPXD4R1 2x Coincidences & 825 & 0.29 & 0.0058  \\
TEPX Clusters & 75 & 0.0915 & 0.00183  \\
TEPX 2x Coincidences in phi & 75 & 0.343 & 0.00685 \\
TEPX 2x Coincidences in r & 75 & 0.593 & 0.0119  \\
TEPX 2x Coincidences & 75 & 0.297 & 0.00594  \\
\end{tabular}
\end{small}
\end{table}

>>>>

The Phase II upgrade of the CMS Tracker is a crucial project undertaken to maintain the effectiveness of the CMS experiment during the High Luminosity LHC (HL-LHC) phase, which is expected to start in 2029. The HL-LHC will significantly increase the instantaneous luminosity, leading to a higher number of proton-proton collisions per bunch crossing. Instantaneous luminosity will increase to an unprecedented value of $7.5 \times 10^{34} cm^{-2} \: s^{-1}$ which corresponds to 200 proton-proton collisions per bunch crossing. This higher collision rate poses several challenges, including increased radiation levels, higher data rates, and more complex event environment. Run 2 pixel detector will not be able to handle these extreme radiation environment, resolve nearby particle tracks and operate properly to give a reliable estimate of the instantaneous luminosity for such a high pileup. That is why it will be replaced by a new pixel detector. The Phase II upgrade of the CMS tracker comprises of two main components: the Inner Tracker (IT) and the Outer Tracker (OT) \cite{collaboration:2759074}. The Inner Tracker consists of three main sub-components: the Tracker Endcap Pixel detector (TEPX), the Tracker Forward Pixel detector (TFPX), and the Tracker Barrel Pixel (TBPX) detector as shown in Fig. \ref{fig:Innertracker} \cite{CERN-LHCC-2017-009}.  %The Outer Tracker is responsible for detecting charged particles in the outer region of the CMS detector. It consists of two key subsystems: the silicon strip tracker and the silicon macro pixel tracker \cite{RoyChowdhury:2729279}.

@@

TEPX Disk 4 Ring 1 (D4R1) is located in the innermost ring of the last double-disk of the TEPX, is uniquely designed for measuring luminosity and beam-induced background (BIB). D4R1 is optimally placed due to its low radius, large active area, and high z position, making it highly sensitive and efficient for these measurements. The D4R1 operational model is distinct from the rest of TEPX \cite{Haranko2023}. It employs the entire available trigger rate and readout bandwidth, which varies based on operational mode (binary readout or Time-over-Threshold information). This capability ranges from a trigger rate of 825 kHz (similar to the rest of TEPX) up to several MHz, limited by readout bandwidth and the available e-link capacity. In nominal physics conditions, the baseline trigger rate for D4R1 is 825 kHz, while in van der Meer (vdM) conditions, it is 2 MHz, with the potential to reach between 5 and 10 MHz in certain scenarios.

@@

Fig. \ref{fig:tepx_cl} shows the location of TEPX clusters in Disk 4 in the XY plane (Z is the direction of the beam) from a simulated sample with a pileup of 200 \cite{dabrowski2020}. %Clusters refer to groups of hits in the pixel detector.
It consists of five concentric rings that are color-coded to indicate the density of clusters within the detector. It is transitioning from blue (low density) to red (high density), with the hottest regions indicating the overlap modules within the rings. The granularity of the data is high, with the number of bins set to 1000 for both axes and a bin size of 0.1 cm, allowing for a detailed visualization of the clusters across the disk.
%The TEPX disk in this map consists of 5 rings which are concentric rings of pixel modules at different radial distances from the beamline.

@@

Simlarly, we also studied two-fold coincidences in r which arise from module overlaps between successive rings by only considering one to one module overlap between first two and last two layers in one TEPX double disk \cite{BRIL2021}.  %Table \ref{table:two_fold_coincidences} categorizes various types of two-fold coincidences in r, observed in the TEPX detector. These coincidences are labeled from A to L and arise due to the module overlap in different layers within a single TEPX double disk. The 'dz' column quantifies the spatial separation along beam axis between the coinciding layers,  the z coordinates for Layers L1 to L4 within the TEPX detector is also mentioned. These coordinates provide a precise location for each layer, contributing to the understanding of where these two fold coincidences are likely to occur. The two fold coincidence in r map showing module overlap regions for different types is shown in Fig. \ref{fig:cluster_ring_770}.
 A one-to-one overlapping of modules is considered between L1L2 and L3L4 module overlap as shown in Fig. \ref{fig:cluster_ring_72034}. %as shown in Fig. \ref{tab:my_label2}. 
 %The distributions for dr, dphi distribution depicting fake, true, total two fold coincidence in r for pileup 0.5 and 200 are shown in Fig. \ref{fig:cluster_ring_81} and \ref{fig:cluster_ring_82} respectively.
 Fit is done to pileup 0.5 distribution of true coincidences for both L1L2 and L3L4 coincidences in r. %as shown in Fig. \ref{fig:cluster_ring_83} and \ref{fig:cluster_ring_84}. 
  Fit results for all disks and rings are shown from  Table \ref{tab:my_label_55} to \ref{tab:my_label_54}. Similarly, two-sigma standard deviation selections are applied on dr, dphi variables. After applying all selections, the XY map of two-fold coincidences in r for L1L2 and L3L4 combined as shown in \ref{fig:cluster_ring_79}.

@@

\begin{table}[htbp]
  \centering
  \caption[Expected stat. precision of TEPX under low pileup] {Expected statistical precision (in $\%$) in head-on collisions during typical vdM conditions with pileup of 0.5 for TEPX clusters and two-fold coincidences over different integration time period \cite{Collaboration:2759074}.}
\begin{tabular}{lcccc}
vdM (PU 0.5) & Trigger Rate (kHz) & 1 bx, 1s & 1 bx, 30s\\
\hline
TEPXD4R1 Clusters&1000&1.82&0.332\\
TEPXD4R1 2x Coincidences in phi &1000&5.98&1.09\\
TEPXD4R1 2x Coincidences in r &1000&11.1&2.02\\
TEPXD4R1 2x Coincidences &1000&5.27&0.962\\
TEPX Clusters&500&0.709&0.129\\
TEPX 2x Coincidences in phi &500&2.65&0.485\\
TEPX 2x Coincidences in r &500&4.59&0.838\\
TEPX 2x Coincidences &500&2.3&0.42\\
\end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\caption[Expected stat. precision of TEPX under high pileup]{Expected statistical precision (in $\%$) in head-on collisions during physics conditions with pileup of 200 for TEPX clusters and two-fold coincidences over different integration time period \cite{Collaboration:2759074}.}
\begin{small} % this command will decrease the font size
\begin{tabular}{lcccc}
Algorithm & Trigger Rate (kHz) & 1 bx, 1s & 2500 bx, 1s\\
\hline
TEPXD4R1 Clusters & 825 & 0.1 & 0.002  \\
TEPXD4R1 2x Coincidences in phi & 825 & 0.329 & 0.00659 \\
TEPXD4R1 2x Coincidences in r & 825 & 0.61 & 0.0122  \\
TEPXD4R1 2x Coincidences & 825 & 0.29 & 0.0058  \\
TEPX Clusters & 75 & 0.0915 & 0.00183  \\
TEPX 2x Coincidences in phi & 75 & 0.343 & 0.00685 \\
TEPX 2x Coincidences in r & 75 & 0.593 & 0.0119  \\
TEPX 2x Coincidences & 75 & 0.297 & 0.00594  \\
\end{tabular}
\end{small}
\end{table}
